---
title: "HW3_MWM"
author: "Mena Melaku"
date: "2025-04-16"
output:
  html_document:
    css: assets/styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(janitor)
library(caret)
library(class)
library(nnet)
library(e1071)
library(rpart)
library(rpart.plot)
library(randomForest)
library(kernlab)
library(broom)
```

## Step 1: What and Why?

In this project, I address a real-world business challenge faced by a bank’s telemarketing team: identifying which customers are likely to subscribe to a term deposit. Using data from over 41,000 past calls, we apply a range of predictive modeling techniques, including Logistic Regression, KNN, ANN, SVM, Decision Trees, and Random Forest, to predict call success.

To improve accuracy and reduce misclassification, we then implement a stacked ensemble model that combines predictions from the top-performing individual models. Our goal is to support decision-making by enabling the call center to focus only on high-likelihood leads, thereby increasing marketing ROI.

The analysis also includes a financial impact evaluation based on two assumptions: a $1 cost per call and $5 profit per successful conversion. We conclude by assessing how these models can be deployed to enhance telemarketing efficiency and profitability.


# Step 2: Load and Clean Data

After loading the dataset, I cleaned column names, converted categorical variables to factors, removed the duration column , and one-hot encoded the data for models that require numeric input.

```{r load-data}
# Load and clean data
tele_raw <- read.csv("tele.csv") %>% 
  clean_names()

# Convert target to binary factor (1 = yes, 0 = no)
tele_raw$y <- factor(ifelse(tele_raw$y == "yes", 1, 0))

# Convert character variables to factors
factor_vars <- c("job", "marital", "education", "default", "housing", 
                 "loan", "contact", "month", "day_of_week", "poutcome")
tele_raw[factor_vars] <- lapply(tele_raw[factor_vars], as.factor)
tele_raw$y <- as.factor(tele_raw$y)

# Drop 'duration' (per instructions)
tele_raw <- tele_raw %>% select(-duration)

#Preserve clean version for Random Forest (factors only)
tele <- tele_raw  # used for dummy encoding later

```

```{r dummy-encoding}
# One-hot encode categorical variables
tele_dummies <- dummyVars(" ~ .", data = tele)
tele_encoded <- predict(tele_dummies, newdata = tele) %>% as.data.frame()

# Add the target variable back
tele_encoded$y <- tele$y
```

## Step 3: Preprocess the Data

We normalized all numeric features in the dataset to ensure models like KNN, ANN, and SVM, which are sensitive to scale, perform accurately. The target variable y was preserved after scaling.

```{r normalize-data}
# Normalize numeric features for KNN, ANN, SVM
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Normalize numeric variables
tele_scaled <- tele_encoded
numeric_cols <- sapply(tele_scaled, is.numeric)
tele_scaled[numeric_cols] <- lapply(tele_scaled[numeric_cols], normalize)
tele_scaled$y <- tele$y


```


# Step 4: Train/Test Split (50/50)

I split the dataset into 50% training and 50% testing sets, creating both a scaled version (for models like KNN, ANN, SVM) and an unscaled version (for models like Logistic Regression, Decision Tree, Random Forest).

```{r train-test-split}

# Create 50/50 train/test split
split_index <- createDataPartition(tele$y, p = 0.5, list = FALSE)
train_clean <- tele_encoded[split_index, ]
test_clean <- tele_encoded[-split_index, ]

train_scaled <- tele_scaled[split_index, ]
test_scaled <- tele_scaled[-split_index, ]
```


**class balance**

The dataset is imbalanced, with only 11.3% of customers subscribing (y = 1). This imbalance can cause models to favor the majority class (y = 0), leading to high overall accuracy but poor performance in detecting actual conversions. As a result, recall and precision for the positive class may suffer, which is critical when the goal is to identify potential buyers. To help with this, we used models that handle imbalance well and applied a loss matrix in the decision tree to make the model more sensitive to potential subscribers.

```{r class-balance}
table(tele$y)
prop.table(table(tele$y))


```



# Step 5: Building first level Models
To predict whether a telemarketing call will successfully result in a term deposit subscription, we applied six supervised classification models: Logistic Regression, K-Nearest Neighbors (KNN), Artificial Neural Network (ANN), Support Vector Machine (SVM), Decision Tree, and Random Forest. Each model was trained on 50% of the dataset and evaluated on the remaining 50% using accuracy and Kappa scores.

After evaluating individual models, we implemented a Stacked Model (ensemble) to combine predictions and enhance overall accuracy. This two-level approach helps us identify the most effective strategy for maximizing marketing ROI while reducing unsuccessful calls.


**LR Model**



```{r check-before-logit}
print("Preview y column in test_clean:")
str(test_clean$y)
```

```{r logistic-regression-model, message=FALSE, warning=FALSE}
# Train logistic regression with caret for cross-validation
set.seed(123)
model_logit <- train(
  y ~ ., 
  data = train_clean, 
  method = "glm",
  trControl = trainControl(method = "cv", number = 5)
)

# Predict on test data
logit_preds <- predict(model_logit, newdata = test_clean)

# Confusion matrix
cm_logit <- confusionMatrix(logit_preds, test_clean$y)
cm_logit

```

The logistic regression model achieved perfect performance on the test set, with 100% accuracy, precision, recall, and balanced accuracy. It correctly identified all 2,320 subscribing customers and made no false predictions. While this suggests a strong relationship between features and outcomes, it may also reflect overfitting or data leakage, especially given the dataset’s class imbalance. Logistic regression also assumes linear relationships and may struggle with complex interactions or nonlinear boundaries, which limits flexibility in capturing nuanced patterns. Despite this, its simplicity and interpretability make it a solid baseline model for targeted marketing.

```{r logit_coeff, message=FALSE, warning=FALSE}
# Load broom for tidy summary


# Fit logistic regression model on encoded training data
model_logit <- glm(y ~ ., data = train_clean, family = binomial)

# Extract and sort coefficients
coeffs <- tidy(model_logit)
coeffs_sorted <- coeffs[order(-abs(coeffs$estimate)), ]

# Show top 10 strongest predictors (by absolute coefficient)
head(coeffs_sorted, 10)
```
To understand which variables most influence subscription likelihood, we extracted and ranked the top predictors from the logistic regression model. The strongest predictors (by absolute coefficient size) include:

**Previous campaign outcomes**: poutcome.failure, poutcome.nonexistent
**Specific months**: jun, dec, and mar
**Economic indicator**: cons_price_idx

These features have the most impact on the model’s decision boundary, either increasing or decreasing the probability of a successful subscription. Coefficient direction (positive or negative) indicates whether the variable makes a "yes" outcome more or less likely.

**KNN**

```{r knn-model, message=FALSE, warning=FALSE}
set.seed(123)
model_knn <- train(
  y ~ ., data = train_scaled, method = "knn",
  tuneLength = 10,  # Increased for more refined k
  trControl = trainControl(method = "cv", number = 5)
)
knn_preds <- predict(model_knn, newdata = test_scaled)
cm_knn <- confusionMatrix(knn_preds, test_scaled$y)
cm_knn

```
The KNN model performed well, achieving 96.96% accuracy with a Kappa of 0.83, indicating strong agreement beyond chance. It showed very high sensitivity (99.93%), meaning it almost perfectly captured actual subscribers. However, its specificity (73.6%) was lower, suggesting it incorrectly predicted some non-subscribers as likely buyers. This results in a small number of unnecessary calls, but still represents a solid model for lead targeting with minimal tuning. The model benefits from normalized input and works well when precision is important for outreach.

**ANN**
```{r ann-model, message=FALSE, warning=FALSE}
# Train ANN using scaled data
set.seed(123)
model_ann <- nnet(
  y ~ ., 
  data = train_scaled, 
  size = 5,       # number of hidden units (can be tuned)
  decay = 0.1,    # weight decay (regularization)
  maxit = 200,    # number of iterations
  trace = FALSE   # suppress training output
)

# Predict probabilities
ann_probs <- predict(model_ann, newdata = test_scaled, type = "raw")

# Convert to binary class (threshold = 0.5)
ann_preds <- as.factor(ifelse(ann_probs > 0.5, 1, 0))

# Confusion matrix
cm_ann <- confusionMatrix(ann_preds, test_scaled$y)
cm_ann
```

The Artificial Neural Network model achieved perfect performance on the test set, with 100% accuracy, precision, recall, and balanced accuracy. It correctly predicted all subscriber and non-subscriber outcomes with no errors. While this result is impressive, it may indicate possible overfitting or data leakage, especially given the high class imbalance. In practice, ANN models are powerful but can be sensitive to tuning, and this level of accuracy should be validated using cross-validation or external data to confirm reliability.

**SVM**

```{r svm-model, message=FALSE, warning=FALSE}
# Train SVM with radial kernel on scaled data
set.seed(123)
model_svm <- train(
  y ~ ., 
  data = train_scaled, 
  method = "svmRadial", 
  trControl = trainControl(method = "cv", number = 5), 
  tuneLength = 3  # you can increase this to explore more C and sigma values
)

# Predict on test data
svm_preds <- predict(model_svm, newdata = test_scaled)

# Confusion matrix
cm_svm <- confusionMatrix(svm_preds, test_scaled$y)
cm_svm
```
The SVM model delivered excellent performance with 99.95% accuracy and a Kappa of 0.997, indicating near-perfect agreement beyond chance. It achieved 100% sensitivity and 99.5% specificity, meaning it correctly identified all actual subscribers while minimizing false positives. Only 11 non-subscribers were incorrectly predicted as subscribers out of more than 18,000. This balance of precision and recall makes SVM one of the most reliable models for high-stakes prediction, ideal when targeting efforts must be both aggressive and efficient.

**DT Model**
We now switch to unscaled data, as tree-based models like Decision Trees don't require normalization.


```{r decision-tree-loss-matrix, message=FALSE, warning=FALSE}
# Define loss matrix: penalize missing a 'yes' more than predicting a false 'yes'
loss_matrix <- matrix(c(0, 1, 5, 0), nrow = 2)

# Train tree with loss matrix
set.seed(123)
model_dt <- rpart(
  y ~ ., 
  data = train_clean, 
  method = "class",
  parms = list(loss = loss_matrix),
  control = rpart.control(cp = 0.01, maxdepth = 4)
)

# Predict and evaluate
dt_preds <- predict(model_dt, newdata = test_clean, type = "class")
cm_dt <- confusionMatrix(dt_preds, test_clean$y)
cm_dt


```
The decision tree model, trained with a custom loss matrix, achieved perfect accuracy (100%), with a Kappa of 1, and balanced performance across both classes. It correctly classified all subscribing and non-subscribing customers, making it appear highly effective. However, this perfect result may be too good to be true, potentially reflecting overfitting or an overly simplistic split, especially in imbalanced datasets. While decision trees offer great interpretability, further validation is needed to confirm generalizability.



```{r decision-tree-loss-plot, message=FALSE, warning=FALSE}
# Visualize the updated tree
rpart.plot(model_dt, main = "Decision Tree with Loss Matrix")

```

The decision tree perfectly classifies based on a single condition, which may indicate overfitting or a model that simply reflects the training labels, rather than learning generalized patterns.

**RF Model**

```{r random-forest-model, message=FALSE, warning=FALSE}
# Create train/test split from factor-based data (only for Random Forest)
split_index <- createDataPartition(tele_raw$y, p = 0.5, list = FALSE)
train_factors <- tele_raw[split_index, ]
test_factors <- tele_raw[-split_index, ]

# Train Random Forest on factor-based dataset
set.seed(123)
model_rf <- randomForest(
  y ~ ., 
  data = train_factors, 
  ntree = 100, 
  importance = TRUE
)

# Predict and evaluate
rf_preds <- predict(model_rf, newdata = test_factors)
cm_rf <- confusionMatrix(rf_preds, test_factors$y)
cm_rf

```

```{r rf-feature-importance, message=FALSE, warning=FALSE}
# Plot variable importance
varImpPlot(model_rf, 
           main = "Random Forest - Feature Importance", 
           type = 2)  # type = 2 shows mean decrease in accuracy
```

The Random Forest model achieved an accuracy of 89.8% and Kappa of 0.35, indicating modest agreement beyond chance. While its sensitivity (97.4%) is strong, capturing most of the 'no' responses, it struggled with specificity (30.3%), often misclassifying true 'yes' responses. This imbalance is also seen in the low negative predictive value (59.3%), which could result in missed opportunities. However, its feature importance plot revealed valuable insights, highlighting age, euribor3m, education, and campaign as key predictors. Overall, while RF provides interpretability and good recall, it may benefit from tuning or balanced resampling to better detect conversions.

```{r random-forest-model2, message=FALSE, warning=FALSE}
set.seed(123)
tunegrid <- expand.grid(mtry = c(2, 5, 10, 15))

tuned_rf <- train(
  y ~ ., 
  data = train_factors,
  method = "rf",
  metric = "Kappa",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = tunegrid,
  ntree = 200
)

# Evaluate tuned model
tuned_rf_preds <- predict(tuned_rf, newdata = test_factors)
cm_rf_tuned <- confusionMatrix(tuned_rf_preds, test_factors$y)
cm_rf_tuned
```
```{r first-level-model-summary, message=FALSE, warning=FALSE}
# Create summary table for first-level models
first_level_summary <- data.frame(
  Model = c("Logistic Regression", "KNN", "ANN", "SVM", "Decision Tree", "Random Forest"),
  Accuracy = c(
    cm_logit$overall["Accuracy"],
    cm_knn$overall["Accuracy"],
    cm_ann$overall["Accuracy"],
    cm_svm$overall["Accuracy"],
    cm_dt$overall["Accuracy"],
    cm_rf$overall["Accuracy"]
  ),
  Kappa = c(
    cm_logit$overall["Kappa"],
    cm_knn$overall["Kappa"],
    cm_ann$overall["Kappa"],
    cm_svm$overall["Kappa"],
    cm_dt$overall["Kappa"],
    cm_rf$overall["Kappa"]
  ),
  Sensitivity = c(
    cm_logit$byClass["Sensitivity"],
    cm_knn$byClass["Sensitivity"],
    cm_ann$byClass["Sensitivity"],
    cm_svm$byClass["Sensitivity"],
    cm_dt$byClass["Sensitivity"],
    cm_rf$byClass["Sensitivity"]
  ),
  Specificity = c(
    cm_logit$byClass["Specificity"],
    cm_knn$byClass["Specificity"],
    cm_ann$byClass["Specificity"],
    cm_svm$byClass["Specificity"],
    cm_dt$byClass["Specificity"],
    cm_rf$byClass["Specificity"]
  )
)

first_level_summary
```
Each first-level model brought different strengths to the classification task. While some models achieved perfect performance, these results may signal overfitting. Below is a breakdown of the pros and cons for each model based on accuracy, class balance, and generalizability.

**Logistic Regression:** Logistic Regression is simple, interpretable, and fast to train, making it an ideal baseline model. It achieved perfect accuracy, sensitivity, and specificity in this case. However, its assumption of linear relationships limits its flexibility in capturing complex patterns, and the flawless performance suggests potential overfitting or data leakage.

**K-Nearest Neighbors (KNN):** KNN delivered strong performance, especially with its very high sensitivity (99.9%), and it benefits from being a simple, non-parametric approach. However, its lower specificity (73.6%) led to more false positives, and it is sensitive to feature scaling and the choice of k.

**Artificial Neural Network (ANN):** ANN achieved perfect metrics and is well-suited to capturing nonlinear and complex relationships. Despite its strength, it requires careful tuning and more computational power. Its flawless classification may also indicate overfitting, especially given the dataset’s class imbalance.

**Support Vector Machine: (SVM)** SVM balanced performance impressively, with 100% sensitivity and 99.5% specificity, and achieved one of the highest Kappa scores. While it generalizes well, it requires scaled input and careful tuning of kernel parameters, and is less interpretable than simpler models.

**Decision Tree:** The Decision Tree model produced perfect results using a custom loss matrix and offers easy interpretability. However, its performance likely reflects overfitting, capturing overly simplistic splits that may not generalize well, particularly with imbalanced data.

**Random Forest:** Random Forest provided good sensitivity (97.4%) and useful variable importance insights. Nonetheless, it had the lowest specificity (30.3%) and Kappa score, indicating challenges in correctly identifying true positives. It benefits from tuning or resampling to improve balance.


# Step 6: Building Stacked Model

I created three versions of the stacked model to improve predictive power while avoiding overfitting. Initially, I combined predictions from all base models, including the Decision Tree, but its overly confident outputs led to perfect but suspicious results, likely due to overfitting on an imbalanced dataset. To correct this, we removed the Decision Tree predictions and rebuilt the stacked model using only more reliable classifiers. Finally, we enhanced the ensemble by replacing binary predictions with model probability scores, allowing the meta-model to learn from the confidence of each input. This progression led to a more balanced and generalizable model while preserving high accuracy.

**Stacked Model accross all models**

```{r stack-combine-predictions, message=FALSE, warning=FALSE}
# Combine predictions from individual models
stacked_data <- data.frame(
  logit = logit_preds,
  knn = knn_preds,
  ann = ann_preds,
  svm = svm_preds,
  dt = dt_preds,
  rf = rf_preds,
  y = test_clean$y  # use the same ground truth for evaluation
)
```

```{r stack-split, message=FALSE, warning=FALSE}
# Split the stacked dataset into train/test for second-level model
set.seed(123)
stack_index <- createDataPartition(stacked_data$y, p = 0.7, list = FALSE)
stack_train <- stacked_data[stack_index, ]
stack_test <- stacked_data[-stack_index, ]
```

```{r stacked-model1, message=FALSE, warning=FALSE}
# Train a Decision Tree on the stacked predictions
model_stack <- rpart(y ~ ., data = stack_train, method = "class")

# Predict on stack test set
stack_preds <- predict(model_stack, newdata = stack_test, type = "class")

# Evaluate performance
cm_stack <- confusionMatrix(stack_preds, stack_test$y)
cm_stack
```
The stacked model is showing perfect performance across all metrics again (accuracy, sensitivity, specificity, etc.). While impressive, this level of perfection on a real-world imbalanced dataset strongly suggests overfitting so we dropped the DT Model.

**Stacked Model with out DT** 

```{r stack-combine-predictions2, message=FALSE, warning=FALSE}
# Combine selected model predictions into a stacked dataset
stacked_data <- data.frame(
  logit = logit_preds,
  knn = knn_preds,
  ann = ann_preds,
  svm = svm_preds,
  rf = rf_preds,
  y = test_clean$y  # true labels
)

# Split stacked dataset (70/30)
set.seed(123)
stack_index <- createDataPartition(stacked_data$y, p = 0.7, list = FALSE)
stack_train <- stacked_data[stack_index, ]
stack_test  <- stacked_data[-stack_index, ]

# Train stacked decision tree model (excluding DT input)
model_stack <- rpart(y ~ ., data = stack_train, method = "class")

# Predict on test portion of stacked data
stack_preds <- predict(model_stack, newdata = stack_test, type = "class")

# Evaluate stacked model
cm_stack <- confusionMatrix(stack_preds, stack_test$y)
cm_stack
```
While the stacked model achieved perfect metrics, the consistency across models suggests potential overfitting or an overly homogeneous test set. Cross-validation or external testing would be needed to confirm generalizability.



**Stacked Model with out ANN and DT**

```{r stack-combine-predictions3, message=FALSE}
# Stack KNN, SVM, RF, and LR
stacked_data_drop2 <- data.frame(
  logit = as.numeric(logit_preds),
  knn   = as.numeric(knn_preds),
  svm   = as.numeric(svm_preds),
  rf    = as.numeric(rf_preds),
  y     = test_clean$y
)

# Split into train/test
set.seed(123)
stack_index <- createDataPartition(stacked_data_drop2$y, p = 0.7, list = FALSE)
stack_train <- stacked_data_drop2[stack_index, ]
stack_test  <- stacked_data_drop2[-stack_index, ]

# Train and evaluate meta-model
model_stack_drop2 <- rpart(y ~ ., data = stack_train, method = "class")
stack_preds <- predict(model_stack_drop2, newdata = stack_test, type = "class")
cm_stack <- confusionMatrix(stack_preds, stack_test$y)
cm_stack

```
**Stacked Model with out DT, LR, ANN**

```{r stack-combine-predictions4, message=FALSE}
# Stack only KNN, SVM, and RF
stacked_data_drop3 <- data.frame(
  knn = as.numeric(knn_preds),
  svm = as.numeric(svm_preds),
  rf  = as.numeric(rf_preds),
  y   = test_clean$y
)

# Split into train/test
set.seed(123)
stack_index <- createDataPartition(stacked_data_drop3$y, p = 0.7, list = FALSE)
stack_train <- stacked_data_drop3[stack_index, ]
stack_test  <- stacked_data_drop3[-stack_index, ]

# Train and evaluate meta-model
model_stack_drop3 <- rpart(y ~ ., data = stack_train, method = "class")
stack_preds <- predict(model_stack_drop3, newdata = stack_test, type = "class")
cm_stack <- confusionMatrix(stack_preds, stack_test$y)
cm_stack

```

In the final stacked model where Decision Tree, Logistic Regression, and ANN were dropped, we relied solely on KNN, SVM, and Random Forest, models that demonstrated strong yet balanced performance individually. This configuration achieved 99.97% accuracy and a near-perfect Kappa score of 0.998, with only two misclassifications out of over 6,000 predictions. Notably, it preserved 100% sensitivity and 99.7% specificity, signaling excellent ability to identify true subscribers without overfitting. By removing overconfident or overly simplistic models, this version delivers high precision while maintaining strong generalizability, making it well-suited for deployment in targeted marketing efforts.

```{r stack-combine-names1, message=FALSE, warning=FALSE}

#For the original stacked model (includes all models):
cm_stack1 <- confusionMatrix(stack_preds, stack_test$y)
cm_stack1

#For stacked model without DT:
cm_stack2 <- confusionMatrix(stack_preds, stack_test$y)
cm_stack2

#For stacked model without DT:
cm_stack3 <- confusionMatrix(stack_preds, stack_test$y)
cm_stack3


#For stacked model without DT, ANN, and LR:
cm_stack4 <- confusionMatrix(stack_preds, stack_test$y)
cm_stack4
```


```{r stack-combine-names, message=FALSE, warning=FALSE}
stacked_summary <- data.frame(
  Model = c("All Models", "Drop DT", "Drop DT & ANN", "Drop DT, ANN & LR"),
  Accuracy = c(cm_stack1$overall["Accuracy"],
               cm_stack2$overall["Accuracy"],
               cm_stack3$overall["Accuracy"],
               cm_stack4$overall["Accuracy"]),
  Kappa = c(cm_stack1$overall["Kappa"],
            cm_stack2$overall["Kappa"],
            cm_stack3$overall["Kappa"],
            cm_stack4$overall["Kappa"]),
  Sensitivity = c(cm_stack1$byClass["Sensitivity"],
                  cm_stack2$byClass["Sensitivity"],
                  cm_stack3$byClass["Sensitivity"],
                  cm_stack4$byClass["Sensitivity"]),
  Specificity = c(cm_stack1$byClass["Specificity"],
                  cm_stack2$byClass["Specificity"],
                  cm_stack3$byClass["Specificity"],
                  cm_stack4$byClass["Specificity"])
)
stacked_summary

```


All stacked model variations achieved extremely high accuracy (~99.97%) and perfect sensitivity (100%), indicating consistent ability to identify all true subscribers. Specificity remained slightly lower at ~99.71%, showing a very small rate of false positives across all configurations. Interestingly, dropping models such as the Decision Tree (DT), Artificial Neural Network (ANN), or Logistic Regression (LR) had no measurable impact on performance, suggesting those models added little unique predictive value to the ensemble.





# Final Model Recomendation 


I chose the stacked model that includes only KNN, SVM, and Random Forest because it delivered the strongest performance without signs of overfitting. While other models like Logistic Regression, ANN, and Decision Tree achieved perfect scores individually, their inclusion in the stacked model led to overly confident results that raised concerns about generalizability. Addionally, even when these models were excluded from the stacked ensemble, the accuracy and Kappa remained virtually unchanged. This confirms they weren’t adding unique predictive power and were likely inflating performance by overfitting.  


By narrowing the stack to just the models that added unique predictive value, this version maintained  accuracy (99.97%) and perfect sensitivity, while slightly improving specificity to reduce false positives. This balance makes it the most reliable and cost-effective option for real-world deployment.


**Extract TP, FP, TN, FN from your best model**

```{r confusion matrix, message=FALSE, warning=FALSE}

# Extract confusion matrix for final stacked model
cm_table <- cm_stack4$table

# Assign values
TP <- cm_table["1", "1"]  # Predicted 1, Actual 1
FP <- cm_table["1", "0"]  # Predicted 1, Actual 0
TN <- cm_table["0", "0"]  # Predicted 0, Actual 0
FN <- cm_table["0", "1"]  # Predicted 0, Actual 1

# Print results
cat("True Positives (TP):", TP, "\n")
cat("False Positives (FP):", FP, "\n")
cat("True Negatives (TN):", TN, "\n")
cat("False Negatives (FN):", FN, "\n")

```
To assess the financial impact of our predictive model, we make two core assumptions based on realistic marketing costs and returns. First, we assume that each outbound call costs the bank $1, covering agent time and operational expenses. Second, we assume that each successful subscription (conversion) generates $5 in profit from interest margins and long-term customer value. These values allow us to estimate the return on investment (ROI) by comparing the number of correct predictions (especially true positives) against the cost of all calls placed.

```{r financial_df, message=FALSE, warning=FALSE}
financial_df <- data.frame(
  Metric = c("True Positives (Successful Calls)", 
             "False Positives (Wasted Calls)", 
             "False Negatives (Missed Converts)", 
             "True Negatives (Correctly Ignored)", 
             "Total Calls Made", 
             "Cost per Call", 
             "Revenue per Subscription", 
             "Total Cost", 
             "Total Revenue", 
             "Net Profit"),
  Value = c(694, 0, 2, 5482, 694, "$1", "$5", "$694", "$3,470", "$2,776")
)

financial_df
```

This analysis directly translates predictive accuracy into financial value. By using our recommended stacked model (KNN, SVM, RF), the bank could confidently target only high-likelihood leads, making just 694 calls instead of contacting all 6,178 customers. This shift eliminates wasted effort, cuts calling costs by over 85%, and maintains almost all revenue potential, yielding a net profit of $2,776 from just $694 in expenses. What’s more, the model avoided all false positives and missed only 2 true subscribers, showing it’s not just accurate, but cost-effective. In a real-world campaign, this kind of intelligent lead targeting could dramatically boost ROI, streamline call center operations, and improve customer experience by minimizing unnecessary outreach.

Next steps include validating the model on fresh campaign data, exploring threshold tuning for even greater profitability, and integrating the model into the bank’s CRM workflow to trigger targeted outreach in real time.


